---
title: "Introduction to simpletable II"
author: "James P. Curley"
date: "Tuesday, April 14, 2015"
output: html_document
---

### Background

The first guide to this package can be [found here](http://rpubs.com/jalapic/simpletable).

<br>
<br>


### Installation

You can install from github with `devtools`:

```{r, eval=FALSE}
library(devtools)
devtools::install_github("jalapic/simpletable")
```


<br>
<br>

### Further Introductory Examples

<br>
<b> Rownames </b>
<br>

The rownames of a dataframe won't automatically be included as the first column.  To get them to show you need to do something like this:

<br>
<br>


```{r, echo=F}
library(simpletable)
```


```{r, results='asis'}
tablecode(cbind(car=rownames(mtcars), mtcars), tabletype="minimal")

```

<br><br>




```{r, results='asis'}
tablecode(cbind(car=rownames(mtcars), mtcars), tabletype="gridtable")
```



<br><br>

### Soccer Example !

Here's an example using data from my [engsoccerdata](https://github.com/jalapic/engsoccerdata).  I'm using a function in that package called `maketable` which converts raw soccer scores into an end of season final league standings table.  I've happened to pick the final standings for the 1994/95 season for the top tier of English soccer. I then turn that into a more attractive table with `tablecode` from simpletable:

<br>


```{r, results='asis', warning=FALSE, message=FALSE}
library(engsoccerdata)
library(dplyr)
engsoccerdata2 %>% 
  filter(Season==1994 & tier==1) %>% 
  maketable() %>%
  as.data.frame() %>%
  tablecode(tabletype="another")

```

<br><br><br>



